---
title: "FCAT Euro 2024 Sweep"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tibble)
library(tidyr)
library(dplyr)
library(here)
library(readr)
library(purrr)
library(reactable)

source(here("R", "league_table.R"))

# seed = round(16.4 * 1224, digits = 0)
now <- "2024-07-15 20:00"
groups <- c("Group A",
            "Group B",
            "Group C",
            "Group D",
            "Group E",
            "Group F")

colspec <- cols(
  matchnumber = col_integer(),
  roundnumber = col_character(),
  date = col_datetime(format = "%d/%m/%Y %H:%M"),
  location = col_character(),
  group = col_character(),
  homeaway = col_character(),
  team = col_character(),
  goals = col_integer(),
  penalties = col_integer(),
  ycards = col_integer(),
  rcards = col_integer(),
  fouls = col_integer()
)

matchdata <- read_csv(here("data", "matchdata.csv"),col_types = colspec) 

fixtures <- matchdata |>  
  pivot_wider(names_from = homeaway,
              values_from = c(team, goals, penalties, ycards, rcards, fouls)) |> 
  mutate(points_hometeam = case_when(goals_hometeam > goals_awayteam ~ 3,
                                     goals_hometeam == goals_awayteam ~ 1,
                                     TRUE ~ 0),
         result_hometeam = case_when(goals_hometeam > goals_awayteam ~ "W",
                                     goals_hometeam == goals_awayteam ~ "D",
                                     TRUE ~ "L"),
         points_awayteam = case_when(goals_awayteam > goals_hometeam ~ 3,
                                     goals_awayteam == goals_hometeam ~ 1,
                                     TRUE ~ 0),
         result_awayteam = case_when(goals_awayteam > goals_hometeam ~ "W",
                                     goals_awayteam == goals_hometeam ~ "D",
                                     TRUE ~ "L"))


hometeam <- fixtures |>  
            select(matchnumber, roundnumber, date, location, group,
                   team = team_hometeam,
                   result = result_hometeam,
                   goals_for = goals_hometeam,
                   goals_against = goals_awayteam,
                   penalties = penalties_hometeam,
                   ycards = ycards_hometeam,
                   rcards = rcards_hometeam,
                   fouls = fouls_hometeam,
                   points = points_hometeam)

awayteam <- fixtures |>  
  select(matchnumber, roundnumber, date, location, group,
         team = team_awayteam,
         result = result_awayteam,
         goals_for = goals_awayteam,
         goals_against = goals_hometeam,
         penalties = penalties_awayteam,
         ycards = ycards_awayteam,
         rcards = rcards_awayteam,
         fouls = fouls_awayteam,
         points = points_awayteam)

matchdata_full <- hometeam |>  
           bind_rows(awayteam) |> 
  filter(date <= now)

```

Welcome to the FCAT Euro 2024 Sweep! Your chance to add a little spice to your
summer and maybe win a valuable Cash Prize! As usual there are a range of
prizes available to maximise the chances of prolonging your interest.

- **Winner** receives £10
- **Runner up** gets £5

In addition there are some bonus categories with a value of £3 each:

- **Wooden Spoon** for the worst team in the tournament based on results
- **Entertainers** for the team involved in the most goals
- **Anti-footballers** for the dirtiest team

All statistics will be drawn from the official
[Euro 2024](https://www.uefa.com/euro2024/) website. The organisers reserve
the right to tweak the criteria for the bonus categories, to break ties or
generally to keep it sensible. As always in case of questions over the rules,
the organisers decision is _final_.

## The Draw

The draw has been conducted in a rigorously reproducible manner, seeded with a
random number. 

Code to reproduce the draw is available
[here](https://github.com/lee269/euro2024).


```{r sweep, echo=FALSE, message=FALSE}



names <- c("Person 1", "Person 2", "Person 3", "Person 4",
           "Person 5", "Person 6", "Person 7", "Person 8", 
           "Person 9", "Person 10", "Person 11", "Person 12", 
           "Person 13", "Person 14", "Person 15", "Person 16",
           "Person 17", "Person 18", "Person 19", "Person 20",
           "Person 21", "Person 22", "Person 23", "Person 24")  

countries <- c("Albania","Austria","Belgium","Croatia",
               "Czechia","Denmark","England","France",
               "Georgia","Germany","Hungary","Italy",
               "Netherlands","Poland","Portugal","Romania",
               "Scotland","Serbia","Slovakia","Slovenia",
               "Spain","Switzerland","Türkiye","Ukraine")

# set.seed(seed)
draw <- tibble(country = sample(countries), name = names) |>  
  arrange(country)

knitr::kable(draw)

```


## The Tournament {.tabset}

Match data as at **`r Sys.time()`**. 

### Fixtures

Use the fixture list to a) to ensure you never miss a match on TV, or b)
schedule some Married At First Sight: Australia binge sessions.

```{r fixtures}

fixtures |> 
  mutate(fixture = ifelse(is.na(goals_hometeam),
                          paste(team_hometeam, "v", team_awayteam),
                          paste(team_hometeam, goals_hometeam, "-", goals_awayteam, team_awayteam)
                          )
         ) |> 
  select(matchnumber, date, fixture, location, group, roundnumber) |> 
  reactable(pagination = FALSE,
            compact = TRUE,
            resizable = TRUE,
            filterable = TRUE,
            columns = list(
            matchnumber = colDef(name = "Match #"),
            date = colDef(name = "Date", format = colFormat(datetime = TRUE)),
            fixture = colDef(name = "Fixture"),
            location = colDef(name = "Venue"),
            group = colDef(name= "Group"),
            roundnumber = colDef(name = "Round"))
            )

```

### Group standings 


```{r groups, echo=FALSE, message=FALSE, results='asis', warning=FALSE}

# league_table <- function(groupname) {
# matchdata_full |> 
#   na.omit() |> 
#   filter(group == groupname) |>
#   group_by(team) |> 
#   summarise(played = n(),
#             goals_for = sum(goals_for),
#             goals_against = sum(goals_against),
#             goals_sum = sum(goals_for) + sum(goals_against),
#             goal_diff = sum(goals_for) - sum(goals_against),
#             ycards = sum(ycards),
#             rcards = sum(rcards),
#             fouls = sum(fouls),
#             thug_index = ((rcards * 25) + (ycards *5) + fouls) / played,
#             pts = sum(points)) |> 
#   # pivot_wider(names_from = result, values_from = played, values_fill = 0) |>
#   # relocate(starts_with(c("W", "L", "D")), .after = team) |>
#   arrange(-pts, -goal_diff) |>  
#     knitr::kable(caption = paste0("**",groupname,"**"))
# }

map(groups, league_table, data = matchdata_full)

```

### Team data

```{r overall}

matchdata_full |> 
  # na.omit() |> 
  filter(!is.na(goals_for)) |> 
  group_by(team) |> 
  summarise(played = n(),
            goals_for = sum(goals_for),
            goals_against = sum(goals_against),
            goals_sum = sum(goals_for) + sum(goals_against),
            goal_diff = sum(goals_for) - sum(goals_against),
            ycards = sum(ycards),
            rcards = sum(rcards),
            fouls = sum(fouls),
            pts = sum(points)) |> 
  arrange(-pts, -goal_diff) |> 
  reactable(pagination = FALSE,
            compact = TRUE)



```


## The Prizes {.tabset .tabset-pills}

### Wooden Spoon

This is based on lowest points total, worst goal difference and most goals
conceded, in that sort order.

```{r wooden-spoon, echo=FALSE}

matchdata_full |> 
  # na.omit() |> 
  filter(!is.na(goals_for)) |> 
  # filter(group %in% groups) |>
  group_by(team) |> 
  summarise(played = n(),
            goals_for = sum(goals_for),
            goals_against = sum(goals_against),
            goals_sum = sum(goals_for) + sum(goals_against),
            goal_diff = sum(goals_for) - sum(goals_against),
            pts = sum(points)) |> 
  arrange(pts, goal_diff, desc(goals_against)) |> 
  reactable(pagination = TRUE,
            compact = TRUE,
            columns = list(
              team = colDef(name = "Team"),
              played = colDef(name = "Matches played"),
              goals_for = colDef(name = "Goals for"),
              goals_against = colDef(name = "Goals against"),
              goals_sum = colDef(name = "Goals total"),
              goal_diff = colDef(name = "Goal difference")
            ) )


```

### Entertainers

The entertainment index is the total goals per game plus goals for per game,
with the tie breaker being the number of goals scored (because we all want to
see some buccaneering all out attacking football). 

``` {r entertainers, echo=FALSE}

matchdata_full |> 
  # na.omit() |> 
  filter(!is.na(goals_for)) |> 
  group_by(team) |> 
  summarise(played = n(),
            goals_for = sum(goals_for),
            goals_against = sum(goals_against),
            goals_sum = sum(goals_for) + sum(goals_against),
            goal_diff = sum(goals_for) - sum(goals_against),
            entertain_index = (goals_sum / played) + (goals_for / played)) |> 
  arrange(-entertain_index, -goals_for) |> 
  reactable(pagination = TRUE, 
            compact = TRUE,
            columns = list(
              team = colDef(name = "Team"),
              played = colDef(name = "Matches played"),
              goals_for = colDef(name = "Goals for"),
              goals_against = colDef(name = "Goals against"),
              goals_sum = colDef(name = "Goals total"),
              goal_diff = colDef(name = "Goal difference"),
              entertain_index = colDef(name = "Entertainment Index", format = colFormat(digits = 2))
            ))



```


### Anti-footballers

The 'thug index' is calculated as follows: a red card is 25pts, a yellow 5pts
and a foul 1. These weights are applied to the match data and divided by the
number of matches played.

```{r thugs, echo=FALSE}

matchdata_full |> 
  # na.omit() |> 
  filter(!is.na(goals_for)) |> 
  group_by(team) |> 
  summarise(played = n(),
            ycards = sum(ycards),
            rcards = sum(rcards),
            fouls = sum(fouls),
            thug_index = ((rcards * 25) + (ycards *5) + fouls) / played) |> 
  arrange(-thug_index) |> 
  reactable(pagination = TRUE,
            compact = TRUE,
            columns = list(
              team = colDef(name = "Team"),
              played = colDef(name = "Matches played"),
              ycards = colDef(name = "Yellow cards"),
              rcards = colDef(name = "Red cards"),
              fouls = colDef(name = "Fouls"),
              thug_index = colDef(name = "Thug index", format = colFormat(digits = 2))
            ))

```


## Payment status

Here is the entrants list and payment status. Thanks for your contribution!.
Please correct me if there are any errors.

```{r payment}

# payment |> arrange(Name, desc(Paid)) |> knitr::kable()


```


## The Winners!


- **Winner** - TBD
- **Runner up** - TBD
- **Wooden Spoon** - TBD - congratulations **`r draw |> filter(country == "Qatar") |> pull(name = )`!**
- **Entertainers** - TBD
- **Anti-footballers** - TBD

